vegan_analysis

### 2024-01-24 scripts/4_vegan_clean.qmd
third submission

bec_normal <- bec_df %>%
  mutate(across(basal_area:VarDHI,  .fns = ~ normal(.x))) 
```

```{r bec-nest}
to_nest <- bec_normal %>%
  select(-percentage_first_returns_above_mean,
         -loreys_height,
         -elev_stddev,
         -elev_mean)

bec_nested <- to_nest %>%
  group_by(bec) %>%
  nest() %>%
  ungroup() %>%
  rename(strata = bec)

class_nested <- to_nest %>%
  group_by(category) %>%
  nest() %>%
  ungroup() %>%
  rename(strata = category)

set.seed(69420)

strata_nested <- bind_rows(bec_nested, class_nested) %>%
  mutate(data = map(data, slice_sample, n = 500))
# this slice sample is the ANALYTICAL difference between 4_vegan and 
# 4_vegan_clean, other than the actual cleanup of the .rmd file

global <- strata_nested %>%
  filter(strata %in% zones) %>%
  unnest(data)
```

@@

rcorr <- summary(rda_out)$sites %>%
  as.data.frame() %>%
  select(starts_with("RDA")) %>%
  bind_cols(global) %>%
  select(-strata,-category) %>%
  as.matrix() %>%
  Hmisc::rcorr()

@@

bec_normal %>%
  pivot_longer(cols = ends_with("DHI"),
               names_to = "dhi",
               values_to = "dhi_val") %>%
  pivot_longer(
    cols = basal_area:total_biomass,
    names_to = "struct",
    values_to = "struct_val"
  ) %>%
  ggplot(aes(x = struct_val, y = dhi_val)) +
  geom_hex() +
  facet_grid(rows = vars(dhi),
             cols = vars(struct),
             scales = "free") +
  labs(x = "Structure Value (Normalized)",
       y = "DHI Value (Normalized)")
```

@@

@@

# list to iterate
stratum <- clean_loads_raw %>%
  pull(strata) %>%
  unique()

# make sure everything gets change in rda fig
possibles <- clean_loads_raw %>%
  select(var, axis) %>%
  complete() %>%
  mutate(possible = paste(var, axis, sep = "-")) %>%
  pull(possible) %>%
  unique()

@@

for (s in stratum) {
  print(s)
  strata_tp <- clean_loads_raw %>%
    filter(strata == s) %>%
    mutate(textreplace = paste(var, axis, sep = "-"),
           axis_loading = as.character(round(axis_loading, digits = 2)))

@@

file.copy(here::here("outputs", "drawio", "British Columbia.drawio"), here::here("outputs", "British Columbia.drawio"), overwrite = T)
```

@@

```{r}
scatter <- bec_normal %>%
  mutate(label = case_when(bec %in% c("CWH", "MS") ~ bec,
                           T ~ "Other"))

@@

>>>>

set.seed(29640)

raw_sliced <- bec_df %>%
  select(-percentage_first_returns_above_mean,
         -loreys_height,
         -elev_stddev,
         -elev_mean) %>%
  group_by(bec) %>%
  slice_sample(n = 500) %>%
  ungroup()

to_nest_normal <- raw_sliced %>% 
  mutate(across(basal_area:VarDHI,  .fns = ~ normal(.x)))
```

```{r bec-nest}
# already slice_sampled
bec_nested <- to_nest_normal %>%
  group_by(bec) %>%
  nest() %>%
  ungroup() %>%
  rename(strata = bec)

# sample down to 500 here
class_nested <- to_nest_normal %>%
  group_by(category) %>%
  slice_sample(n = 500) %>%
  nest() %>%
  ungroup() %>%
  rename(strata = category)

strata_nested <- bind_rows(bec_nested, class_nested)
# this slice sample is the ANALYTICAL difference between 4_vegan and 
# 4_vegan_clean, other than the actual cleanup of the .rmd file

global <- to_nest_normal
```

@@

rcorr <- summary(rda_out)$sites %>%
  as.data.frame() %>%
  select(starts_with("RDA")) %>%
  bind_cols(global) %>%
  select(-bec,-category) %>%
  as.matrix() %>%
  Hmisc::rcorr()

@@

@@

```{r}
egg <- anova.cca(rda_out, step = 1000)

outs %>%
  mutate(anova_sig = map(.x = rda, .f = \(x) {
    anova.cca(rda, step = 1000)
  })
```

@@

# list to iterate
stratum <- clean_loads_cum_pos %>%
  pull(strata) %>%
  unique()

# make sure everything gets change in rda fig
possibles <- clean_loads_cum_pos %>%
  select(var, axis) %>%
  complete() %>%
  mutate(possible = paste(var, axis, sep = "-")) %>%
  pull(possible) %>%
  unique()

@@

for (s in stratum) {
  print(s)
  strata_tp <- clean_loads_cum_pos %>%
    filter(strata == s) %>%
    mutate(textreplace = paste(var, axis, sep = "-"),
           axis_loading = as.character(round(axis_loading, digits = 2)))

@@

# generate and run batch script to export all files to png, then copy the main
# one to the main folder

in_drawio <- list.files(here::here("outputs", "drawio"), full.names = T) %>%
  str_subset("_rda", negate = T)

out_drawio <- in_drawio %>%
  str_replace("\\.drawio", ".png")

bat_loc <- here::here("scripts", "export_drawio.bat")

glue::glue('"C:/Program Files/draw.io/draw.io.exe" -x "{in_drawio}" -o "{out_drawio}"') %>%
  write_lines(bat_loc)

system2(bat_loc)

file.copy(here::here("outputs", "drawio", "British Columbia.png"), here::here("outputs", "British Columbia.drawio.png"), overwrite = T)
```

@@

```{r}
scatter <- raw_sliced %>%
  mutate(label = case_when(bec %in% c("CWH", "MS") ~ bec,
                           T ~ "Other"))

@@

Matrix of plots as requested by reviewer #4

Density hexplots of forest structural attributes and the Dynamic Habitat Indices. Note the log scale.

```{r}
(hex_matrix_plot <- raw_sliced %>%
  pivot_longer(cols = basal_area:total_biomass, names_to = "struct", values_to = "struct_vals") %>%
  pivot_longer(ends_with("DHI"), names_to = "DHI", values_to = "dhi_vals") %>%
  left_join(keys$continuous, by = c("struct" = "variable")) %>%
  mutate(struct_vals = struct_vals / divide_by,
         unit = ifelse(struct == "elev_cv", "", glue::glue("({unit})")),
         label = glue::glue("{var_long} \n{unit}")) %>%
  ggplot(aes(x = struct_vals, y = dhi_vals)) +
  geom_hex(aes(fill = after_stat(log10(count)))) +
  facet_grid(rows = vars(DHI),
             cols = vars(label),
             scales = "free",
             switch = "both") +
  labs(x = NULL,
       y = NULL) +
  scale_fill_continuous_c4a_seq(palette = "viridis") +
  theme(legend.position = "bottom",
        strip.placement = "outside",
        strip.background = element_blank())

)

ggsave(here::here("outputs", "hex_matrix_plot.png"),
       height = 6, width = 10)
```

to_nest_normal %>%
  pivot_longer(cols = ends_with("DHI"),
               names_to = "dhi",
               values_to = "dhi_val") %>%
  pivot_longer(
    cols = basal_area:total_biomass,
    names_to = "struct",
    values_to = "struct_val"
  ) %>%
  left_join(keys$continuous, by = c("struct" = "variable")) %>%
  mutate(struct_vals = struct_val / divide_by,
         unit = ifelse(struct == "elev_cv", "", glue::glue("({unit})")),
         label = glue::glue("{var_long} \n{unit}")) %>%
  ggplot(aes(x = struct_val, y = dhi_val)) +
  geom_hex(aes(fill = after_stat(log10(count)))) +
  facet_grid(rows = vars(dhi),
             cols = vars(var_long),
             scales = "free",
             switch = "both") +
  labs(x = "Structure Values (Normalized)",
       y = "DHI Values (Normalized)") +
  scale_fill_continuous_c4a_seq(palette = "viridis") +
  theme(legend.position = "bottom",
        strip.placement = "outside",
        strip.background = element_blank())
```
